---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Autocorrelação Espacial e Aprendizado de máquina

*Laís de Souza Teixeira & Alan Rodrigo Panosso* 


```{r,message=FALSE,warning=FALSE,error=FALSE}
# devtools::install_github("Envirometrix/plotKML")
# devtools::install_github("Envirometrix/landmap")
library(tidyverse)
library(sp)
library(geobr)
library(skimr)
library(GSIF)
library(geoR)
library(raster)
library(ranger)
```

```{r}
dados <- read_rds("data/dados_lais.rds")
skim(dados)
```

# Seguindo o exemplo de Meuse

```{r}
data("meuse")
```

## Ajustando o modelo


```{r}
zinc.geo <- as.geodata(meuse[c("x","y","zinc")])
ini.v <- c(var(log1p(zinc.geo$data)),500)
zinc.vgm <- likfit(zinc.geo, lambda=0, ini=ini.v, cov.model=
"exponential")
```


```{r}
data("meuse.grid")
data("meuse")
locs <- as.geodata(meuse.grid)$coords
meuse.grid <- SpatialPointsDataFrame(meuse.grid[1:2],meuse.grid)
meuse.grid <- SpatialPointsDataFrame(meuse[1:2],meuse)
zinc.ok <- krige.conv(zinc.geo, locations=locs, krige=krige.control
(obj.m=zinc.vgm))
```



## Buffer distance
which derives a gridded map for each observation point. The spatial prediction model is defined as:

```{r}
data("meuse.grid")
data("meuse")
grid.dist0 <- GSIF::buffer.dist(
  SpatialPointsDataFrame(meuse[1:2],meuse),  
  SpatialPixelsDataFrame(points = meuse.grid[c("x", "y")], data = meuse.grid)[1],
  as.factor(1:nrow(meuse))
)
```


```{r}
dn0 <- paste(names(grid.dist0), collapse="+")
fm0 <- as.formula(paste("zinc ~ ", dn0))
```


which means that the target variable is a function of 155 covariates. Next, we overlay points and covariates to create a regression matrix, so that we can tune and fit a ranger model, and generate predictions:

```{r}
data("meuse")
ov.zinc <- over(SpatialPointsDataFrame(meuse[1:2],meuse), grid.dist0)
rm.zinc <- cbind(meuse["zinc"], ov.zinc)
m.zinc <- ranger(fm0, rm.zinc, quantreg=TRUE, num.trees=150, 
                 mtry = 98,
                 min.node.size = 4)
m.zinc
```


```{r}
zinc.rfd <- predict(m.zinc, grid.dist0@data)
str(zinc.rfd)
```


```{r}
meuse.grid$zinc_rfd = zinc.rfd$predictions
meuse.grid$zinc_ok = zinc.ok$predict
#meuse.grid$zinc_rfd_range = (zinc.rfd[,3]-zinc.rfd[,1])/2
cor.test(zinc.rfd$predictions,zinc.ok$predict)
```

```{r}
as.tibble(meuse.grid) |> 
  ggplot(aes(x=x,y=y)) +
  geom_tile(aes(fill = zinc_ok)) +
  ggplot2::scale_fill_gradient(low = "yellow", high = "blue") +
  ggplot2::coord_equal()
```


```{r}
as.tibble(meuse.grid) |> 
  ggplot(aes(x=x,y=y)) +
  geom_tile(aes(fill = zinc_rfd)) +
  ggplot2::scale_fill_gradient(low = "yellow", high = "blue") +
  ggplot2::coord_equal()
```

