---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Autocorrelação Espacial e Aprendizado de máquina

*Laís de Souza Teixeira & Alan Rodrigo Panosso* 


```{r,message=FALSE,warning=FALSE,error=FALSE}
# devtools::install_github("Envirometrix/plotKML")
# devtools::install_github("Envirometrix/landmap")
library(tidyverse)
library(sp)
library(geobr)
library(skimr)
library(GSIF)
library(geoR)
library(raster)
library(ranger)
```

```{r}
dados <- read_rds("../data/dados_lais.rds")
skim(dados)
```

# Dados

```{r}
da <- dados |> filter(data == "2015-10-02") |> 
  mutate(FCO2 = ifelse(FCO2 <0, mean(FCO2),FCO2) ) |> 
  dplyr::select(x,y,FCO2)
```

## Ajustando o modelo


```{r}
fco2.geo <- as.geodata(da[c("x","y","FCO2")])
ini.v <- c(var(log1p(fco2.geo$data)),500)
fco2.vgm <- likfit(zinc.geo, lambda=0, ini=ini.v, cov.model=
"exponential")
```


```{r}
dis <- 2
da.grid <- expand.grid(x=seq(min(da$x),max(da$x),dis), y=seq(min(da$y),max(da$y),dis))
da.grid$part.a<-0
locs <- as.geodata(da.grid)$coords
da.grid <- SpatialPointsDataFrame(da[1:2],da)
fco2.ok <- krige.conv(fco2.geo, locations=locs, krige=krige.control
(obj.m=fco2.vgm))
```



## Buffer distance
which derives a gridded map for each observation point. The spatial prediction model is defined as:

```{r}
dis <- 2
da.grid <- expand.grid(x=seq(min(da$x),max(da$x),dis), y=seq(min(da$y),max(da$y),dis))
da.grid$part.a<-0
grid.dist0 <- GSIF::buffer.dist(
  SpatialPointsDataFrame(da[1:2],da),  
  SpatialPixelsDataFrame(points = da.grid[c("x", "y")], data = da.grid)[1],
  as.factor(1:nrow(da))
)
```


```{r}
dn0 <- paste(names(grid.dist0), collapse="+")
fm0 <- as.formula(paste("FCO2 ~ ", dn0))
```


which means that the target variable is a function of 155 covariates. Next, we overlay points and covariates to create a regression matrix, so that we can tune and fit a ranger model, and generate predictions:

```{r}
ov.fco2 <- over(SpatialPointsDataFrame(da[1:2],da), grid.dist0)
rm.fco2 <- cbind(da["FCO2"], ov.fco2)
m.fco2 <- ranger(fm0, rm.fco2, quantreg=TRUE, num.trees=1500, 
                 mtry = 100,
                 min.node.size = 5)
m.fco2
```


```{r}
fco2.rfd <- predict(m.fco2, grid.dist0@data)
str(fco2.rfd)
```


```{r}
da.grid$FCO2_rfd = fco2.rfd$predictions
da.grid$FCO2_ok = fco2.ok$predict
#meuse.grid$zinc_rfd_range = (zinc.rfd[,3]-zinc.rfd[,1])/2
cor.test(zinc.rfd$predictions,zinc.ok$predict)
```

```{r}
as.tibble(da.grid) |> 
  ggplot(aes(x=x,y=y)) +
  geom_tile(aes(fill = FCO2_ok)) +
  ggplot2::scale_fill_gradient(low = "yellow", high = "blue") +
  ggplot2::coord_equal()
```


```{r}
as.tibble(da.grid) |> 
  ggplot(aes(x=x,y=y)) +
  geom_tile(aes(fill = FCO2_rfd)) +
  ggplot2::scale_fill_gradient(low = "yellow", high = "blue") +
  ggplot2::coord_equal()
```

